Evaluation of 3-point correlation functions from structural images on CPU and GPU architectures: accounting for anisotropy effects

### 2024-04-23 paper.tex
Renaming

\subsection{Computational algorithms}
Three-point correlation function $S_3$ can be computed either in spatial domain
or in frequency domain via the convolution theorem. Suppose that a finite
countable set $A$ is defined as a multidimensional array. A point $x$ belongs to
a subset $A_n$ if and only if $A[x] = n$. The simplest algorithm below
calculates $S_3$ pointwise in the spatial domain:
\begin{algorithmic}[1]
  \Procedure{$S_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \Comment Apply $I_n$ to the input array elementwise.
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A''_n \gets \mathfrak{S}_{x_2}(A_n)$
  \State $T \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
In this algorithm the indicator function $I_n$ for the phase of interest $n$ is
applied to the input array $A$. Then the shift operator $\mathfrak{S}$ is used
to shift $A_n$ by vectors $x_1$ and $x_2$. Finally, the three arrays are
multiplied pointwise and a normalized sum of all elements of the product is
returned.

@@

Algorithms for three-point surface functions are the same as for $S_3$,
containing only small modifications. The main difference is that we need to
replace an indicator function $I_n$ with an edge detection operator $E_n$
(see Appendix \ref{sec:filter}). Then the algorithms are straightforward:
\begin{algorithmic}[1]
  \Procedure{$F_{sss}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A''_n \gets \mathfrak{S}_{x_2}(A_n)$
  \State $T \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{ssv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A'_{(void)} \gets \mathfrak{S}_{x_2}(A_{(void)})$
  \State $T \gets A_n \cdot A'_n \cdot A'_{(void)}$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{svv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_{(void)} \gets \mathfrak{S}_{x_1}(A_{(void)})$
  \State $A''_{(void)} \gets \mathfrak{S}_{x_2}(A_{(void)})$
  \State $T \gets A_n \cdot A'_{(void)} \cdot A''_{(void)}$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}

@@

To compute $C_3$ function we need to use a connected-component labeling
algorithm \cite{4728561,PhysRevB.14.3438} to label clusters. Then, again, a
slightly modified algorithm for $S_3$ is used to assess cluster function:
\begin{algorithmic}[1]
  \Procedure{$C_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \State $C_n \gets \mathfrak{C}(A_n)$
  \State $C'_n \gets \mathfrak{S}_{x_1}(C_n)$
  \State $C''_n \gets \mathfrak{S}_{x_2}(C_n)$
  \State $T \gets C_n \odot C'_n \odot C''_n$
  \State \textbf{return} $Sum(T) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
Here $\mathfrak{C}$ is connected-component labeling operator and $\odot$ is
defined as follows:
\begin{equation}
  x \odot y = \left\{
  \begin{array}{ll}
    1 & \quad x = y \\
    0 & \quad \text{otherwise}.
  \end{array}
  \right.
\end{equation}

@@

To test $F_{sss}$ function we use a recently developed approach which computes
that function for sets defined by inequality $f(x) \le T$, $x \in \mathbb{R}^3$
with help of automatic differentiation \cite{postnicov20232}. We choose $f$ as:
\begin{equation}
  f(x, y, z) = y^2 + z^2 - a x^2 (b^2 - x^2)
\end{equation}
and parameters $a = 10$, $b = 0.45$, $T = 10^{-2}$, so inequation
$f(x, y, z) \le T$ gives us a dumbbell-like object (\cref{fig:sss-dumbbell}).

@@

To estimate the error introduced by rotation, consider a
$300 \times 300 \times 300$ sample of value noise \cref{fig:roterror-noise}
which is defined as $f(x, y, z) > T$ where $f$ is a value noise function and $T$
is a threshold. We produce images of the noise rotated around $Z$ axis by $\phi$
radians by discretizing an inequation $f(p(x'), p(y'), p(z')) > T$ where
\begin{equation*}
  \begin{aligned}
    x' &= x \cos \phi + y \sin \phi \\
    y' &= y \cos \phi - x \sin \phi \\
    z' &= z
  \end{aligned}
\end{equation*}
and
\begin{equation*}
  \begin{aligned}
    p(x) &= \left\{
    \begin{array}{ll}
      x'     & \quad x' < 1 \\
      2 - x' & \quad \text{otherwise}
    \end{array}
    \right. \\
    x ' &= x \mod 2
  \end{aligned}
\end{equation*}
Therefore we assume the noise to be periodic with the period equal to
$2$ and the interface between phases is continuous. Discretizing for different
values of $\phi \in [0, \pi/4]$ we obtain the first dataset which contains the
best possible discretization of the rotated noise.

>>>>

\subsection{Computational algorithms}
Three-point correlation function $S_3$ can be computed either in spatial domain
or in frequency domain via the convolution theorem. Suppose that a finite
countable set $A$ is defined as a multidimensional array. A point $x$ belongs to
a subset $A_n$ if and only if $A[x] = n$. The simplest algorithm below
calculates $S_3$ pointwise in the spatial domain:
\begin{algorithmic}[1]
  \Procedure{$S_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \Comment Apply $I_n$ to the input array elementwise.
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A''_n \gets \mathfrak{S}_{x_2}(A_n)$
  \State $Tmp \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(Tmp) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
In this algorithm the indicator function $I_n$ for the phase of interest $n$ is
applied to the input array $A$. Then the shift operator $\mathfrak{S}$ is used
to shift $A_n$ by vectors $x_1$ and $x_2$. Finally, the three arrays are
multiplied pointwise and a normalized sum of all elements of the product is
returned.

@@

Algorithms for three-point surface functions are the same as for $S_3$,
containing only small modifications. The main difference is that we need to
replace an indicator function $I_n$ with an edge detection operator $E_n$
(see Appendix \ref{sec:filter}). Then the algorithms are straightforward:
\begin{algorithmic}[1]
  \Procedure{$F_{sss}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A''_n \gets \mathfrak{S}_{x_2}(A_n)$
  \State $Tmp \gets A_n \cdot A'_n \cdot A''_n$
  \State \textbf{return} $Sum(Tmp) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{ssv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_n \gets \mathfrak{S}_{x_1}(A_n)$
  \State $A'_{(void)} \gets \mathfrak{S}_{x_2}(A_{(void)})$
  \State $Tmp \gets A_n \cdot A'_n \cdot A'_{(void)}$
  \State \textbf{return} $Sum(Tmp) / Norm(A, x_1, x_2)$
  \EndProcedure

  \Procedure{$F_{svv}$}{$A, n, x_1, x_2$}
  \State $A_n \gets E_n (A)$
  \State $A_{(void)} \gets I_{(void)} (A)$
  \State $A'_{(void)} \gets \mathfrak{S}_{x_1}(A_{(void)})$
  \State $A''_{(void)} \gets \mathfrak{S}_{x_2}(A_{(void)})$
  \State $Tmp \gets A_n \cdot A'_{(void)} \cdot A''_{(void)}$
  \State \textbf{return} $Sum(Tmp) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}

@@

To compute $C_3$ function we need to use a connected-component labeling
algorithm \cite{4728561,PhysRevB.14.3438} to label clusters. Then, again, a
slightly modified algorithm for $S_3$ is used to assess cluster function:
\begin{algorithmic}[1]
  \Procedure{$C_3$}{$A, n, x_1, x_2$}
  \State $A_n \gets I_n (A)$
  \State $C_n \gets \mathfrak{C}(A_n)$
  \State $C'_n \gets \mathfrak{S}_{x_1}(C_n)$
  \State $C''_n \gets \mathfrak{S}_{x_2}(C_n)$
  \State $Tmp \gets C_n \odot C'_n \odot C''_n$
  \State \textbf{return} $Sum(Tmp) / Norm(A, x_1, x_2)$
  \EndProcedure
\end{algorithmic}
Here $\mathfrak{C}$ is connected-component labeling operator and $\odot$ is
defined as follows:
\begin{equation}
  x \odot y = \left\{
  \begin{array}{ll}
    1 & \quad x = y \\
    0 & \quad \text{otherwise}.
  \end{array}
  \right.
\end{equation}

@@

To test $F_{sss}$ function we use a recently developed approach which computes
that function for sets defined by inequality $f(x) \le U$, $x \in \mathbb{R}^3$
with help of automatic differentiation \cite{postnicov20232}. We choose $f$ as:
\begin{equation}
  f(x, y, z) = y^2 + z^2 - a x^2 (b^2 - x^2)
\end{equation}
and parameters $a = 10$, $b = 0.45$, $U = 10^{-2}$, so inequation
$f(x, y, z) \le U$ gives us a dumbbell-like object (\cref{fig:sss-dumbbell}).

@@

To estimate the error introduced by rotation, consider a
$300 \times 300 \times 300$ sample of value noise \cref{fig:roterror-noise}
which is defined as $f(x, y, z) > U$ where $f$ is a value noise function and $U$
is a threshold. We produce images of the noise rotated around $Z$ axis by $\phi$
radians by discretizing an inequation $f(p(x'), p(y'), p(z')) > U$ where
\begin{equation*}
  \begin{aligned}
    x' &= x \cos \phi + y \sin \phi \\
    y' &= y \cos \phi - x \sin \phi \\
    z' &= z
  \end{aligned}
\end{equation*}
and
\begin{equation*}
  \begin{aligned}
    p(x) &= \left\{
    \begin{array}{ll}
      x'     & \quad x' < 1 \\
      2 - x' & \quad \text{otherwise}
    \end{array}
    \right. \\
    x ' &= x \mod 2
  \end{aligned}
\end{equation*}
Therefore we assume the noise to be periodic with the period equal to
$2$ and the interface between phases is continuous. Discretizing for different
values of $\phi \in [0, \pi/4]$ we obtain the first dataset which contains the
best possible discretization of the rotated noise.
